---
title: "Practical Machine Learning Course Project"
output: html_document
---

This course project uses accelerometer data from 6 participants performing barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har. We will use machine learning to build a model that can predict the 5 cases based on the accelerometer data. We will then use the model to predict for a test data set wether the excercises were performed correctly or incorrectly.

# Download data
```{r Get data}
setwd("C:/Users/deuscoli/Desktop/Coursera_DataScience/8_Practical_Machine_Learning/")
fileUrl1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileUrl2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl1, destfile = "pml-training.csv", method = "wininet")    
download.file(fileUrl2, destfile = "pml-testing.csv", method = "wininet")
```


# Data preprocessing
Next we do some data cleaning and subset the all the training data into a training and validation set so that we can validate our model later. The data set has 160 variables. We need to retain the most important ones.
```{r Preprocessing}
library(caret)
testing <- read.csv("pml-testing.csv")
all_train <- read.csv("pml-training.csv")
inTrainSet <- createDataPartition(all_train$classe, p = 0.8, list = FALSE)
training <- all_train[inTrainSet, ]
validation <- all_train[-inTrainSet, ]

classe <- training$classe

# Remove index variable x
training <- training[ , -training$X]

# Identify and remove columns with NAs
training<-training[,colSums(is.na(training)) == 0]

# Identify and remove variables with near zero variance as they will not have any predictive power.
columnIndex <- nearZeroVar(training)
training <- training[ , -columnIndex]
#validation <- validation[ , -columnIndex]

# Identify and remove non numeric variables
numIndex <- sapply(training, is.numeric)
training <-cbind(classe,training[,numIndex])
```

# Build Model
We build a model using random forests.
```{r Build and validate model, cache = TRUE}
set.seed(1313)
modelFit <- train(training$classe ~ . , method="rf", data=training)
modelFit$finalModel
```


# Validate Model
Use model to predict the validation data set and do confusion matrix to check performance.
```{r Validate mode}
validation <- validation[ , names(training)] # Limit validation set to only columns in training set
vclasse <- validation$classe
validation <- validation[ , -1]

predv <- predict(modelFit$finalModel, validation)
print(confusionMatrix(predv, vclasse))
```

The accuracy of our model is 0.9995 when tested against the validation data set.

# Predict test data set
Now we can predict the outcomes for our 20 test cases.
```{r Predict test set}
# Limit test set to only columns in training set
testing <- testing[ , names(training)[-1]]
predict(modelFit$finalModel, testing)
```
